A user-extensible and adaptable parser architecture

### 2008-05-29 doc/logparser-acronyms.tex
Changed RBL to DNSBL, and added a reference for it.

\newacronym{RBL}{Real-time Black List}{description={
    Real-time Black Lists are a simple collaborative anti-spam technique
    used to reject or penalise email sent from mail servers reported to
    have sent large volumes of spam.  To use an RBL the mail server makes a
    DNS request incorporating the IP address of the currently connected
    client; if the requested hostname exists the client is on the RBL, and
    the mail server can decide what course of action to take.  XXX
    REFERENCES XXX.
}}

>>>>

\newacronym{DNSBL}{DNS Blacklist}{description={
    A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
    anti-spam technique used to reject or penalise email sent from mail
    servers reported to have sent large volumes of spam.  To use a DNSBL
    the mail server makes a DNS request incorporating the IP address of the
    currently connected client; if the requested hostname exists the client
    is on the DNSBL, and the mail server can decide what course of action
    to take.
}}

### 2008-05-29 doc/logparser-background.tex
Misc notes and minor changes.
Changed RBL to DNSBL, and added a reference for it.

The restriction \texttt{check\_client\_access} checks whether the \IP{}
address of the connected client is found in the file and returned the
associated action if found; the method of searching the file is dependant
on the type of the file (\texttt{cidr} in the example) --- see
\cite{postfix-lookup-tables} for more details.  Other restrictions
determine their result by consulting external sources, e.g.\
\texttt{reject\_rbl\_client rbl.example.com} checks whether the \IP{}
address of the client is present in the \RBL{} \texttt{rbl.example.com},
rejecting the command if the client is listed.
This description is necessarily brief, for further details
see~\cite{smtpd_access_readme, smtpd_per_user_control, policy-servers}.

@@

>>>>

The restriction \texttt{check\_client\_access} checks whether the \IP{}
address of the connected client is found in the file and returned the
associated action if found; the method of searching the file is dependant
on the type of the file (\texttt{cidr} in the example) --- see
\cite{postfix-lookup-tables} for more details.  Other restrictions
determine their result by consulting external sources, e.g.\
\texttt{reject\_rbl\_client dnsbl.example.com} checks whether the \IP{}
address of the client is present in the \DNSBL{}
\texttt{dnsbl.example.com}, rejecting the command if the client is listed.
This description is necessarily brief, for further details
see~\cite{smtpd_access_readme, smtpd_per_user_control, policy-servers}.

@@

XXX CHANGE THIS; THE PARSER DESIGN IS NOVEL, DON'T STRESS THE SIMILARITY
TOO MUCH\@.

### 2008-05-29 doc/logparser-coverage.tex
Changed RBL to DNSBL, and added a reference for it.

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILES{} contiguous log files from Postfix 2.2 and 2.3), and new
rules are easy to add.  Maintaining full coverage is hard because other
servers have different restrictions with custom messages, \RBL{} messages
change over time, major releases of Postfix change warning messages
(usually adding more information), etc.,\ so over time the log lines drift
and change.  Graph~\refwithpage{rule hits graph} shows the number of hits
for each rule over all \numberOFlogFILES{} log files; it's obvious that a
small number of rules match the vast majority of the lines, and more than
half the rules match fewer than 100 times.

>>>>

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILES{} contiguous log files from Postfix 2.2 and 2.3), and new
rules are easy to add.  Maintaining full coverage is hard because other
servers have different restrictions with custom messages, \DNSBL{} messages
change over time, major releases of Postfix change warning messages
(usually adding more information), etc.,\ so over time the log lines drift
and change.  Graph~\refwithpage{rule hits graph} shows the number of hits
for each rule over all \numberOFlogFILES{} log files; it's obvious that a
small number of rules match the vast majority of the lines, and more than
half the rules match fewer than 100 times.

### 2008-05-29 doc/logparser-introduction.tex
Changed RBL to DNSBL, and added a reference for it.

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check whether new anti-spam measures are working.  The
more adventurous will have generated statistics to show how many hits each
of their anti-spam measures has gotten in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this type of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\RBL{}\footnote{This document is supplied with a glossary, see
appendix~\ref{Glossary}.} returns a different explanatory message; policy
servers may log different messages depending on the characteristics of the
connection; there are many ways in which the log lines may differ between
servers, even within the same organisation --- servers may be configured
differently, or running different versions of Postfix.  This paper
documents the difficult process of parsing Postfix log files, presenting
\PLP{}, a program which parses log files and places the resulting data into
a database for later use.  The gathered data can then be used to optimise
current anti-spam measures, provide a baseline to test new anti-spam
measures against, or to produce statistics showing how effective those
measures are.  There are numerous other uses for such data: improving
server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks which propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

>>>>

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check whether new anti-spam measures are working.  The
more adventurous will have generated statistics to show how many hits each
of their anti-spam measures has gotten in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this type of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\DNSBL{}\footnote{This document is supplied with a glossary, see
appendix~\ref{Glossary}.} returns a different explanatory message; policy
servers may log different messages depending on the characteristics of the
connection; there are many ways in which the log lines may differ between
servers, even within the same organisation --- servers may be configured
differently, or running different versions of Postfix.  This paper
documents the difficult process of parsing Postfix log files, presenting
\PLP{}, a program which parses log files and places the resulting data into
a database for later use.  The gathered data can then be used to optimise
current anti-spam measures, provide a baseline to test new anti-spam
measures against, or to produce statistics showing how effective those
measures are.  There are numerous other uses for such data: improving
server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks which propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

### 2008-05-29 doc/logparser-other-parsers-review.tex
Changed RBL to DNSBL, and added a reference for it.

\url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline Fields
extracted: from, to, \RBL{}, client hostname and client \IP{}\@.

\url{http://www.thesawmill.co.uk/formats/beta_postfix.html} \newline Fields
extracted: from, to, client hostname, client \IP{}, relay hostname, relay
\IP{}, status, response code, \RBL{} and message id.

Even if the three data sets were linked together Sawmill would extract less
data than \parsername{}, and it does not appear to extract data about
rejections except when the rejection is caused by an \RBL{} check.

>>>>

\url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline Fields
extracted: from, to, RBL, client hostname and client \IP{}\@.

\url{http://www.thesawmill.co.uk/formats/beta_postfix.html} \newline Fields
extracted: from, to, client hostname, client \IP{}, relay hostname, relay
\IP{}, status, response code, RBL and message id.

Even if the three data sets were linked together Sawmill would extract less
data than \parsername{}, and it does not appear to extract data about
rejections except when the rejection is caused by an \DNSBL{} check.

### 2008-05-29 doc/logparser-parsing-rules.tex
Changed RBL to DNSBL, and added a reference for it.

The log files produced by Postfix differ from installation to installation,
because administrators have the freedom to choose the subset of available
restrictions which suits their needs, including using different \RBL{}
services, policy servers, or custom rejection messages.  To facilitate easy
parsing of new log lines, the parser's design separates parsing rules from
parsing actions: adding new actions is difficult, but adding new rules to
parse new rejection messages is trivial (and also occurs much more
frequently).  The implementation provides a program to ease the process of
creating new rules from unparsed log lines, based on the algorithm
developed by Risto Vaarandi~\cite{risto-vaarandi} for his
\SLCT{}~\cite{slct-paper}.  The differences between the two algorithms will
be outlined as part of the general explanation below.

>>>>

The log files produced by Postfix differ from installation to installation,
because administrators have the freedom to choose the subset of available
restrictions which suits their needs, including using different \DNSBL{}
services, policy servers, or custom rejection messages.  To facilitate easy
parsing of new log lines, the parser's design separates parsing rules from
parsing actions: adding new actions is difficult, but adding new rules to
parse new rejection messages is trivial (and also occurs much more
frequently).  The implementation provides a program to ease the process of
creating new rules from unparsed log lines, based on the algorithm
developed by Risto Vaarandi~\cite{risto-vaarandi} for his
\SLCT{}~\cite{slct-paper}.  The differences between the two algorithms will
be outlined as part of the general explanation below.

### 2008-05-29 papers/sgai-2008/paper.tex
Checked every URL in the bibliography.
Changed RBL to DNSBL, and added a reference for it.

XXX CHECK EVERY URL IN THE BIBLIOGRAPHY\@.

@@

    \item External resources utilised by Postfix (e.g.\ policy
        servers~\cite{policy-servers} or \RBL{} lookups) can also change
        their messages, sometimes without warning.

@@

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejections), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type, and only that type (e.g.\ rejection by a specific
\RBLshort{}).

@@

>>>>

XXX IN THE INTRODUCTION\@: WHY IS THIS RELEVANT TO AN AI CONFERENCE\@?

XXX MAYBE MENTION THAT LOG LINES COULD BE MODIFIED AND RE-PARSED\@?  MAYBE
WORK IT INTO THE ARCHITECTURE CHARACTERISTICS SECTION, AND ELSEWHERE\@.

XXX MAYBE ADD AN EXAMPLE RULE AND LOG LINE, IF THERE'S SUFFICIENT SPACE\@.

@@

    \item External resources utilised by Postfix (e.g.\ policy
        servers~\cite{policy-servers} or \DNSBL{} lookups) can also change
        their messages, sometimes without warning.

@@

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejections), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type, and only that type (e.g.\ rejection by a specific
\DNSBL{}).

@@

System could be used by mail administrators to monitor restrictions.

Explain how someone could build upon it, e.g.\ to parse further Postfix
stuff, or to parse other logs entirely.

``This is a real world application, based on a real world need, rather than
a toy application built as a proof of concept''  Put this in the
introduction as well.
